The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-11-21 manuscript/manuscript.qmd

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

@@

@@

TODO: Andrew/Meng - overview of ONGO law and multiple pathways - TAs vs. ROs - take lots of the stuff from the introduction and move it here

@@

@@

## Closing space in China

>>>>

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The laws requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China.As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

TODO: Andrew - update and coordinate the introductory paragraphs with the theoretical framing below

*Meng's notes/suggestions*: to highlight the theoretical significance in the very beginning (just listing the idea pool for moment, not concise enough):

 1. A frequently approach to close the space of INGOs is passing restrictive laws/institutional approach. [@DeMattee:2019]. To further understanding this global phenomena , China is an important case to explore the impacts such restrictive laws have on the operation of INGOs. Such legal restrictions can be either direct (denial of legal access/cannot register) or covert (red-tape, requiring the approval of PSUs, or the uncertainty of having to filing TAs case by case)

 2. *de facto* vs. *de jure* (incorporating the paragraph below) and differentiated treatment for more vs. less contentious INGOs. Authoritarian -> arbitrary and varied application of the law -> different *operating space*

@@

3. Laws and policies as a *policy tool* and behavioral shifit: to tame INGO to be more cooperative to government. The requirement of PSU and Chinese partner to re-direct INGOs to be more authority friendly.

The impacts on international civil society  we examine include not only access or representation of INGOs in China, but also whether there is change of behaviors [@WuNie:2021]. Mention the concept of *flexibility (in the sense of adjusting to change)* too?

@@

TODO: Meng - add brief explanations of the TA/RO pathways figure and articulate the "cooperative requirements part", i.e., seeking a psu or Chinese partner

@@

(Meng added)

@@

## China's INGO as a policy tool 
*note* changed the sub-title to incorporate the policy tool lense too.
